Reading Notes

### 2020-09-09 2pd.tex
preliminary work

Ultimately the goal this research is to help solve the global energy crisis. 
Solar energy is an important part of a green energy solution.
This research aims to gain insight about new OPV compound morphologies using molecular simulation.
In order to achieve this goal, we have defined the following objectives:
\begin{enumerate}
    \item {Develop a generalized toolset for coarse-graining and performing simulations of novel molecules for organic photovoltaics}
    \item {Develop a general tool to calculate atomistic positions from a coarse-grained morphology}
    \item {Predict nanostructures of novel OPV chemistries}
    \item {Identify chemical features correlated with robust self-assembly and good charge transport}
\end{enumerate}

@@

To improve the useability and extensibility of molecular simulation tools, I have been intentional about incorporating Software Carpentry priniciples into my work.
Background for thinking about users
pedagogy for thinking about how users take in new information
useability
I have completed Software Carpentry training which has taught me how to break down complex ideas and give learners a chance to use their knowledge.
With these principles in mind, I have created tutorial examples to help current and future students understand scientific concepts and how to use computational tools\cite{notebooktutorials}

In my scientific software development with the Molecular Simulation and Design Framework (MoSDeF) team
I have contributed to development of open-source packages including mbuild, foyer, signac, and fresnel\cite{mbuild, foyer, signac, fresnel}.
These contributions have given me ample opportunity to excercise best software development practices like writing good documentation, using version control, writing unit tests, employing CI, using a fork \& pull workflow, etc.
This experience will help me to be more efficient in my future work.
I've also gained valuable experience as working as part of a team.
As I plan to integrate my future work with existing projects, good collaborative skills will be necessary.
growth -- nervousness to participate or submit prs empower and shape the way i think about my thesis

A specific example of what we mean by improving the transferrability, reproducibility, useability, and extensibility is the coarse-graining tools that I have developed on route to fufilling Objective 1 of this proposal.
As for work towards my proposed plan, I have created a minimum working example of a coarse-graining and backmapping package which uses SMARTS strings to specify the site chemistries\cite{grits}.
This package will need further testing as it is tested on different compounds, but the proof of concept exists.

Packages which do coarse-graining and backmapping already exist, so why is another one necessary?
One such package, Versatile Object-oriented Toolkit for Coarse-graining Applications (VOTCA)\cite{Ruhle2011b}, includes a coarse-graining, forcefield creation with reverse MC or IBI, and back-mapping modules.
It also allows for charge transport calculations using KMC similar to MorphCT\cite{Lukyanov2010}.
VOTCA makes some progress towards a TRUE simulation pipeline: they include a docker container with their pre-built software stack and example input files.
However, some of these input files don't work and the unit test coverage on their examples repo at the of time of this writing is 0\%.
While including example inputs and a prebuilt software stack via docker is a great first step, these tutorials give very little walkthrough aside from code comments buried in the xml\cite{votca-github}.
\begin{wrapfigure}{l}{0.5\linewidth}
    \centering
    \includegraphics[width=\linewidth]{images/hexane-compare.pdf}
    \caption{Hexane chemical structure (top) overlaid with coarse-grained mapping \textcolor{blue}{A}-\textcolor{orange}{B}-\textcolor{blue}{A} (bottom)}\label{hexane}
\end{wrapfigure}
The code is a mix of C++, shell commands, and python, and the user is required to interact with compiled programs through the command-line which makes adding this code into an existing workflow a little more difficult.
Overall though this project has many features, interfaces with many popular MD engines (LAMMPS, GROMACS, HOOMD), and is under active, collaborative development, it is near-overwhelming in its scope and is not designed for the new user.
By choosing a pure python API I will be able to make use of Jupyter Notebooks which can be run on cloud servers using tools like MyBinder which can combine executable code with markdown text and images\cite{mybinder, jupyter, notebooktutorials}.
In this way a tutorial can provide exact commands needed to reproduce the example intimately connected with the rationale and justification of that example.

For an example of what I mean by cognitive load, let's compare an example of my proposed coarse-graining scheme to VOTCAs.
In the following example we will take a hexane molecule and map it to three coarse-grained sites (two carbon atoms per site, see~\autoref{hexane}).

VOTCA uses an xml file to do this:

@@

It may be easy to think that because the goal is simple, mapping two carbons and their attached hydrogens to a single site so that six alkyl carbons becomes three coarse-grained sites (A-B-A), that the process will also be simple.
In order to specify a site-mapping using VOTCA, we need to know the residue number, the residue name, and the atom name.
For a six atom system this is not too bad, but it could get tedious and error-prone for larger systems.
And keep in mind that these xml files need to be created manually for each compound which is coarse grained or  backmapped.
VOTCA's scheme also requires that the bonding and the angles resulting from those bonds are specified, another process which is tedious, error-prone, and probably unecessary. 
To explain, let's discuss my proposed alternative:
\begin{lstlisting}
hexane_sites = [("_A", "C[CH3]"), ("_B", "CC")]
\end{lstlisting}
The result of using this code in the \lstinline{coarse} function is a three-site hexane model A-B-A, same as VOTCA\cite{grits}.
The mapping here is specified as a list of site name SMARTS string pairs.
Although they do add some element of cognitive load, SMARTS strings are designed to be human and machine readable and can match atoms based on their chemical environment.
With this scheme, there is no need to look up the name or number of each residue, I simply specify that want two alkyl carbons (in SMARTS notation a capital \lstinline{"C"} denotes an alkyl carbon) and requiring that one be a terminal carbon by including that it has three attached hydrogens (\lstinline{"[CH3]"}).
The order of the sites here matters because technically all the carbon atoms in hexane could be mapped with the more general SMARTS string (\lstinline{"CC"}), so it is important to list the most specific strings first.
The bonding and angles of the resulting coarse-grained sites can then inferred based on the bonding of the mapped atoms.
Basically what VOTCA's xml file does in 51 lines, I can do in 1 which is easier to read and makes chemical sense. 
There are even benefits to the mapping being inline with the code:
It allows the user to see the mapping as they progress through the workflow without having to search for multiple files.
Not using an external file may prevent common errors caused by manual file IO such as operating system specific line endings or whitespace characters like tab or space being read differently by a program.
Having the mapping in-line with the workflow also allows for the user to check that the mapping is correct with built in interactive visualizations, which can display the mapping as an overlay. 
(\autoref{hexane} is the output of one such visualization using py3Dmol\cite{py3dmol}.)

We propose a way to make the mapping and backmapping process easier for the user.
Simplified Molecular Input Line Entry System or SMILES was designed to represent organic molecules in an unambiguous way (similar to IUPAC) but in a line-entry system which is also readable by a computer\cite{Weininger1988}.
This makes it a good candidate for specifying a coarse-grain mapping scheme.
The open-source package, Open Babel, allows for a SMILES string to be converted to a chemical structure and vice versa\cite{openbabel, OBoyle2011}.
SMARTS is similar to SMILES but instead of representing a single molecule unambiguously, it has multiple wildcards which can be used for pattern matching\cite{daylight}.
This pattern matching is also implemented in the Open Babel package.
If we know the SMARTS string for any given CG mapping, we can map the CG site to its center of mass or geometric center.
I have developed a minimum working example of how this can be done and then how the atomistic structure can be recovered\cite{grits}.
This mapping scheme could even be used to make the mapping process easier for established coarse-grained models.
It could be fairly straightforward to implement a lookup to check if site parameters already exist (e.g., for the MARTINI model) by going through and writing out the SMARTS pattern for each existing site type.
This could make parameterizing new compounds for an existing method more foolproof.

>>>>

Ultimately the goal this research is to help solve the global energy crisis. 
Solar energy is an important part of a green energy solution.
This research aims to gain insight about new OPV compound morphologies using molecular simulation.
In order to achieve this goal, we have defined the following objectives:
\begin{enumerate}
    \item {Develop a generalized toolset for coarse-graining and performing simulations of novel molecules for organic photovoltaics}
    \item {Develop a general tool to calculate atomistic positions from a coarse-grained morphology}
    \item {Predict nanostructures of novel OPV chemistries}
    \item {Identify chemical features correlated with robust self-assembly}
    \item {Identify chemical features correlated with good charge transport}
\end{enumerate}

@@

To improve the useability and extensibility of molecular simulation tools, I have been intentional about incorporating Software Carpentry priniciples into my work.
My experience as a Software Carpentry instructor has given me the pedagocial foundation to think about how users take in new information.
With these principles in mind, I have created tutorial examples to help current and future students understand scientific concepts and how to use computational tools\cite{notebooktutorials}

An area where transferrability, reproducibility, useability, and extensibility is really brought to the forefront is in my experience developing scientific software with the Molecular Simulation and Design Framework (MoSDeF) team.
I have contributed to development of open-source packages including mbuild, foyer, signac, and fresnel\cite{mbuild, foyer, signac, fresnel}.
These contributions have given me ample opportunity to excercise best software development practices like writing good documentation, using version control, writing unit tests, employing CI, using a fork \& pull workflow, etc.
This experience will help me to be more efficient in my future work.
I've also gained valuable experience as working as part of a team.
As I plan to integrate my future work with existing projects, good collaborative skills will be necessary.
I have also found this experience to be very empowering: submitting changes to reputable codebases has helped me to place myself in the computational science community.
Helping others to realise this feeling for themselves is part of the reason I so strongly believe in the principles of TRUE science.

A specific example of what I mean by improving the transferrability, reproducibility, useability, and extensibility is the coarse-graining tools that I have developed on route to fufilling Objective 1 of this proposal.
I have created a minimum working example of a coarse-graining and backmapping package which uses SMARTS strings to specify the site chemistries\cite{grits}.
Many packages which do coarse-graining and backmapping already exist, so it is worth asking why is another one necessary?
To asnwer, let's look at one such package, Versatile Object-oriented Toolkit for Coarse-graining Applications (VOTCA)\cite{Ruhle2011b}, which includes coarse-graining, forcefield creation with reverse MC or IBI, and back-mapping modules.
It also allows for charge transport calculations using KMC similar to MorphCT\cite{Lukyanov2010}.
VOTCA makes some progress towards a TRUE simulation pipeline: they include a docker container with their pre-built software stack and example input files.
However, some of these input files don't work and the unit test coverage on their examples repo at the of time of this writing is 0\%.
While including example inputs and a prebuilt software stack via docker is a great first step, these tutorials give very little walkthrough aside from code comments buried in the xml\cite{votca-github}.

\begin{wrapfigure}{l}{0.5\linewidth}
    \centering
    \includegraphics[width=\linewidth]{images/hexane-compare.pdf}
    \caption{Hexane chemical structure (top) overlaid with coarse-grained mapping \textcolor{blue}{A}-\textcolor{orange}{B}-\textcolor{blue}{A} (bottom)}\label{hexane}
\end{wrapfigure}
The code is a mix of C++, shell commands, and python, and the user is required to interact with compiled programs through the command-line which makes adding this code into an existing workflow a little more difficult.
Overall, although this project has many features, interfaces with many popular MD engines (LAMMPS, GROMACS, HOOMD), and is under active, collaborative development, it is near-overwhelming in its scope and is not designed for the new user.

For an example of what I mean by cognitive load, let's compare my proposed coarse-graining scheme to VOTCA's.
In the following example we will take a hexane molecule and map it to three coarse-grained sites (two carbon atoms per site, see~\autoref{hexane}).
VOTCA uses an xml file to do this:

@@

The goal is simple---mapping two carbons and their attached hydrogens to a single site, so six alkyl carbons becomes three coarse-grained sites (A-B-A).
However, in order to specify a mapping using VOTCA, we need to know the residue number, the residue name, and the atom name.
For a six atom system this is not too bad, but it could get tedious and error-prone for larger systems.
And, keep in mind, these xml files need to be created manually for each compound which is coarse grained or backmapped.
VOTCA's scheme also requires that the resulting bonds and angles are specified, another process which is tedious, error-prone, and probably unecessary. 
To explain, let's discuss my proposed alternative:
\begin{lstlisting}
hexane_sites = [("_A", "C[CH3]"), ("_B", "CC")]
\end{lstlisting}
The result of using this code in the \lstinline{coarse} function is a three-site hexane model A-B-A, same as VOTCA\cite{grits}.
The mapping here is specified as a list containing pairs of site name and SMARTS string.
Although they do add some cognitive load, SMARTS strings are designed to be human and machine readable and can be used to match atoms based on their chemical environment.
Simplified Molecular Input Line Entry System or SMILES was designed to represent organic molecules in an unambiguous way (similar to IUPAC) but in a line-entry system which is also readable by a computer\cite{Weininger1988}.
SMARTS is similar to SMILES but instead of representing a single molecule unambiguously, it has multiple wildcards which can be used for pattern matching\cite{daylight}.
This makes it a good candidate for specifying a coarse-grain mapping scheme.
The open-source package, Open Babel, allows for a SMILES string to be converted to a chemical structure and for chemical pattern matching\cite{openbabel, OBoyle2011}.
If we know the SMARTS string for any given CG mapping, we can map the CG site to its center of mass or geometric center.
With this scheme, there is no need to look up the name or number of each residue, a user simply specifies that they want two alkyl carbons (in SMARTS notation a capital \lstinline{"C"} denotes an alkyl carbon) and requiring that one be a terminal carbon by including that it has three attached hydrogens (\lstinline{"[CH3]"}).
The order of the sites here matters because technically all the carbon atoms in hexane could be mapped with the more general SMARTS string (\lstinline{"CC"}), so it is important to list the most specific strings first.
The bonding and angles of the resulting coarse-grained sites can then inferred based on the bonding of the mapped atoms.
Basically what VOTCA's xml file does in 51 lines, I can do in 1 which is easier to read and makes chemical sense.

Beyond just the mapping operator being shorter and easier to understand, there are other ways this codebase considers TRUE principles in its development.
By choosing a pure python API I will be able to make use of Jupyter Notebooks which can be run on cloud servers using tools like MyBinder which can combine executable code with markdown text and images\cite{mybinder, jupyter, notebooktutorials}.
In this way a tutorial can provide exact commands needed to reproduce the example intimately connected with the rationale and justification of that example.
The mapping operator being inline with the code allows the user to see the mapping as they progress through the workflow without having to search for multiple files.
Not using an external file may prevent common errors caused by manual file IO such as operating system specific line endings or whitespace characters like tab or space being read differently by a program.
Having the mapping in-line with the workflow also allows for the user to check that the mapping is correct with built in interactive visualizations, which can display the mapping as an overlay. 
(\autoref{hexane} is the output of one such visualization using py3Dmol\cite{py3dmol}.)
By keeping these principles in mind throughout development we can make things better. %% help
